Rewind: Automatically Generated\\Digital Memories from Mobile Geolocation Data

### 2016-09-21 paper/uist_2016/paper.tex
about to work on memorability findings

Our research questions examine the effectiveness of Rewinds compared to static photos through three virtues of a memory: memorability, believability, and desirability. % heather - do we need a reference for the three virtues? april - yeah that would be ideal!

% heather - I changed this part...
\begin{enumerate}% TODO TODO define semantic vs episodic memory
\item Memorability: How well do Rewinds facilitate the retrieval of semantic and episodic memory? What types of participants are most likely to recall their memories using Rewinds?
\item Believability:  How accurate are Rewinds in the eyes of the participants? 
\item Desirability: What factors are inherent to participants' decisions in preserving artifacts created by the Rewind system?
\end{enumerate}

\subsection{Participants}
To recruit a diverse local population, participants for the user study were recruited through online forums (Craigslist classifieds and local Reddit communities) and offline ads, which were posted in cafes and other small businesses in the the city. Of the 21 total participants, 13 identified themselves as men and 8 identified themselves as women, with ages ranging from 20 to 50 years old. Participants were mainly local residents, with occupations including librarian, medical researcher, office worker, IT consultant, student, musician, and comedian.
% Jeff notes, comment this out because it's not remarkable
% A fifth of the participants were undergraduate or graduate students.

@@

At the beginning of the study, participants were asked to identify their gender, whether they considered themselves a frequent traveler, and whether they tended to use GPS in places they were already familiar with.
Then, questions about the photos and Rewinds investigated their memorability, believability, and desirability.

%Participants who used GPS in familiar locations were considered regular GPS users, and participants who did not were considered non-regular GPS users.
For each photo-Rewind pair, the participant was first shown a photo of one of the locations found in their geolocation data. The participant was asked to answer the following questions about the photo: Do you remember this location? If so, can you name the street? Can you name what is around the location, such as buildings, landmarks, or other places? Would you want to keep this photo? Why or why not? Were you traveling alone? If you remember the location, is the photo accurate to your memory? Why or why not? What was different?

% heather - Rewind of the trip or Rewind video of the trip? april - we say 'rewind' when we mean rewind video, and if we want to refer to the system, we say 'the rewind system'
The participant was then shown a Rewind of the trip that they were on when the location in the photo was encountered. While the Rewind was playing, the participant could hover over the Rewind panel and scrub backwards and forwards horizontally through the images that composed the video. This allowed participants to ``walk through'' their memory at their own pace. The participant was asked the following questions while the Rewind played: Do you remember these locations? Do you remember this particular trip? Can you name some of the streets? Can you name what is around these locations, such as buildings, landmarks, or other places? Would you want to keep this video? Why or why not? Were you traveling alone? If you remember these locations, is the video accurate to your memory? Why or why not? What was different?

@@

% heather - ...that they did not carry the device on their person. - grammatical error? april - nope (maybe out of fashion now though?), their person = their body
Data from three participants were not included in the results of this study. One of the participants' data points consisted of only one location. The participant used a mobile device provided by the study, so it is likely that they did not carry the device on their person throughout the day, but left it in one place. Another participant did not complete the in-person interview portion of the user study due to health and scheduling reasons. The last participant only collected three days of data because they had their phone replaced during the tracking period, and did not reactivate Google's location reporting services on their new device. This participant completed the in-person interview portion of the user study, but we did not include the data gathered during this participant's session in the results.

% heather - need to be rephrased perhaps? Should we give a high level summary of all the relationships we've found? More analysis to be done for photo vs. videos, another table?
% semantic vs episodic memory
% mention distance categories in the beginning (up up up) + everything else 
\section{Findings}
The findings of our user study center on the participants’ relationship to the Rewinds generated from their location data compared to the static photos. We focus on the participants’ judgments of their Rewinds’ memorability, believability, and desirability in comparison to their judgments of the photos paired with the Rewinds. We also discuss the participants' interactions with and reactions to their Rewinds. The results use data from the user studies conducted with 12 men and 6 women. Half of the participants identified themselves as frequent travelers. Additionally, half of the participants were regular GPS users, which we defined as people who use GPS navigation even in locations already familiar to them.

% heather - where to put this part or do we need a summary like this at all?
% april - moved to findings section from user study section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Our user study explored the following five metrics as potentially having an effect on memorability, believability, or desirability of digital memory media:
\begin{enumerate}
	\item Distance: how does the location's distance away from the participant's home influence the participant's perception of Rewinds and static photos? Distances ranged from 0 to 206,155 meters. For analysis, these distances were logarithmically scaled and normalized into six distance groups to mimic human categorization of distances into generalizations like ``close,'' ``far,'' or ``very far.'' When normalized: Group 0 corresponded to a range of 0--10 meters from the participant's home location; Group 1 corresponded to a range of 10--100 meters; Group 2 corresponded to a range of 100--1,000 meters; Group 3 corresponded to a range of 1,000--10,000 meters; Group 4 corresponded to a range of 10,000--100,000 meters and Group 5 corresponded to a range of 100,000--1,000,000 meters from the participant's home location.  
	\item Visit frequency: do participants react differently to Rewinds and static photos depicting locations they visit frequently compared to locations they rarely visit? We showed locations that were either ``most frequently visited,'' ``least frequently visited,'' or randomly selected.
	\item GPS usage: do participants who use GPS navigation regularly react differently to Rewinds and static photos compared to participants who don't use GPS navigation regularly?
	\item Traveler type: do attitudes toward Rewinds and static photos differ depending on whether the participant identifies as a frequent traveler?
	\item Companions: how does the presence of a traveling companion affect the participant's attitude toward Rewinds or static photos?
\end{enumerate}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Within-subject study}
\subsubsection{Semantic Memory} 
Within-subject comparison in our study reveals that participants are significantly better at recalling semantic memory of a particular location with \textit{Rewind} videos than with static street-view images, $\chi^2(1, N=356) = 96.987$, $p <.01$. When being presented with a street-side image retrieved based on their location data, participants were able to recognize places visited 50\% of the time; yet recognition rate of locations reached 97\% when participants started playing accompanied Rewinds, with a significant increase of 47\%. It is perhaps not too surprising as cinematic media like Rewinds would provide users with more contextual meanings and retrival cues for semantic processing, as three participants (P4, P6 and P10) have once expressed their contentment after watching Rewinds by claiming ``I recognize the photo now!'' when their initial attempts with static images recognition were fruitless.

\subsection{Between-Subject Study}
%%UPDATE BOTH TABLES AND ADD ONE MORE%%%
%%%%%%%%%%%percentage to decimal
\begin{table*}[t]
	\centering
	\caption{My caption}
	\small
	\label{within-subject-table}
	\begin{tabular}{c|c|c|c|c|c|c|c|c|}
		\cline{2-9}
		& \multicolumn{4}{c|}{Memorability} & \multicolumn{2}{c|}{Belivability} & \multicolumn{2}{c|}{Desirability} \\ \cline{2-9} 
		& \begin{tabular}[c]{@{}c@{}}video\\  location\end{tabular} & \multicolumn{2}{c|}{\begin{tabular}[c]{@{}c@{}}photo\\ location/trip\end{tabular}} & \begin{tabular}[c]{@{}c@{}}video\\ trip\end{tabular} & photo & video & photo & video \\ \hline
		\multicolumn{1}{|c|}{N} & 177 & \multicolumn{2}{c|}{179} & 177 & 86 & 128 & 179 & 177 \\ \hline
		\multicolumn{1}{|c|}{\%} & 96.61\% & \multicolumn{2}{c|}{49.72\%} & 53.67\% & 77.91\% & 79.69\% & 10.61\% & 14.69\% \\ \hline
		\multicolumn{1}{|c|}{p-val} & \multicolumn{2}{c|}{\begin{tabular}[c]{@{}c@{}}$\chi^2(1, 356) = 96.987$\\ p = 0.0001\end{tabular}} & \multicolumn{2}{c|}{\begin{tabular}[c]{@{}c@{}}$\chi^2(1, 356) = 0.41$\\ p = 0.2611\end{tabular}} & \multicolumn{2}{c|}{\begin{tabular}[c]{@{}c@{}}$\chi^2(1, 214) = 0.02$\\ p = 0.4434\end{tabular}} & \multicolumn{2}{c|}{\begin{tabular}[c]{@{}c@{}}$\chi^2(1, 356 ) = 0.995$\\ p = 0.1593\end{tabular}} \\ \hline
	\end{tabular}
\end{table*}
% 2nd table %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

% shall we change photo/rewind video rememberance to location/trip rememberance? april - i define photo/rewind remembrance below. a photo and a rewind can have both location and trip remembrance. for instance from a photo you might remember the location but not the trip. 
\subsection{Memorability}
Rewinds, and digitally-mediated memories in general, are meant to engage a person's memory. To find out what features make a digital memory memorable, we examined participant responses to whether or not they were able to remember the location shown in a photo (semantic memory) or recall the trip portrayed by a Rewind (episodic memory). Photo memorability is determined by whether or not the participant remembered the location depicted in a photo. Because a Rewind contains many locations, it would be unfair to consider the memorability of Rewind based on whether the participant remembers any locations in the Rewind. So, Rewind memorability is determined by whether or not the participant remembered and could distinguish the particular trip portrayed in a Rewind. In this section, we will look at differences in photo and Rewind memorability based on the five metrics discussed previously.

>>>>

Our research questions examine the effectiveness of Rewinds compared to static photos through three virtues of a memory: memorability, believability, and desirability. % TODO cite why we choose these three topics

\begin{enumerate}
\item Memorability: How well do Rewinds facilitate the retrieval of a memory? What types of participants are most likely to recall their memories using Rewinds?
\item Believability:  How accurate are Rewinds in the eyes of the participants? 
\item Desirability: What factors are inherent to participants' decisions in preserving artifacts created by the Rewind system?
\end{enumerate}

\subsection{Participants}
To recruit a diverse local population, participants for the user study were recruited through online forums on Craigslist classifieds and local Reddit communities, and through offline ads, which were posted in cafes and other small businesses in the the city. Of the 21 total participants, 13 identified themselves as men and 8 identified themselves as women, with ages ranging from 20 to 50 years old. Participants were mainly local residents, with occupations including librarian, medical researcher, office worker, IT consultant, student, musician, and comedian.
% A fifth of the participants were undergraduate or graduate students.

@@

At the beginning of the study, participants were asked to identify their gender, whether they considered themselves a frequent traveler, and whether they tended to use GPS in places they were already familiar with. Then, questions were asked about photos and Rewinds to evaluate their memorability, believability, and desirability.

For each photo-Rewind pair, the participant was first shown a photo of one of the locations found in their geolocation data. The participant was asked to answer the following questions about the photo: Do you remember this location? If so, can you name the street? Can you name what is around the location, such as buildings, landmarks, or other places? Would you want to keep this photo? Why or why not? Were you traveling alone? If you remember the location, is the photo accurate to your memory? Why or why not? What was different?

The participant was then shown a Rewind of the trip that they were on when the location in the photo was encountered. While the Rewind was playing, the participant could hover over the Rewind panel and scrub backwards and forwards horizontally through the images that composed the video. This allowed participants to ``walk through'' their memory at their own pace. The participant was asked the following questions while the Rewind played: Do you remember these locations? Do you remember this particular trip? Can you name some of the streets? Can you name what is around these locations, such as buildings, landmarks, or other places? Would you want to keep this video? Why or why not? Were you traveling alone? If you remember these locations, is the video accurate to your memory? Why or why not? What was different?

@@

Data from three participants were not included in the results of this study. One of the participants' data points consisted of only one location. The participant used a mobile device provided by the study, so it is likely that they did not carry the device on their person throughout the day, but left it in one place. Another participant did not complete the in-person interview portion of the user study due to health and scheduling reasons. The last participant only collected three days of data because they had their phone replaced during the tracking period, and did not reactivate Google's location reporting services on their new device. This participant completed the in-person interview portion of the user study, but we did not include the data gathered during this participant's session in the results.

% heather - need to be rephrased perhaps? Should we give a high level summary of all the relationships we've found? More analysis to be done for photo vs. videos, another table?
% semantic vs episodic memory
% mention distance categories in the beginning (up up up) + everything else 
\section{Findings}
The findings of our user study center on the participants’ relationships to the Rewinds generated from their location data compared to the static photos. We focus on the participants’ judgments of their Rewinds’ memorability, believability, and desirability in comparison to their judgments of the photos paired with the Rewinds. We also discuss the participants' interactions with and reactions to their Rewinds. The results use data from the user studies conducted with 12 men and 6 women. Half of the participants identified themselves as frequent travelers. Additionally, half of the participants were regular GPS users, which we defined as people who use GPS navigation even in locations already familiar to them.

Our user study explored the following five metrics as potentially having an effect on memorability, believability, or desirability of digital memory media:
\begin{enumerate}
	\item Distance: how does the location's distance away from the participant's home influence the participant's perception of Rewinds and static photos? Distances ranged from 0 to 206,155 meters. For analysis, these distances were logarithmically scaled and normalized into six distance groups to mimic human categorization of distances into generalizations like ``close,'' ``far,'' or ``very far.'' When normalized: Group 0 corresponded to a range of 0--10 meters from the participant's home location; Group 1 corresponded to a range of 10--100 meters; Group 2 corresponded to a range of 100--1,000 meters; Group 3 corresponded to a range of 1,000--10,000 meters; Group 4 corresponded to a range of 10,000--100,000 meters and Group 5 corresponded to a range of 100,000--1,000,000 meters from the participant's home location.  
	\item Visit frequency: do participants react differently to Rewinds and static photos depicting locations they visit frequently compared to locations they rarely visit? We showed locations that were either ``most frequently visited,'' ``least frequently visited,'' or randomly selected.
	\item Companions: how does the presence of a traveling companion affect the participant's attitude toward Rewinds or static photos?
	\item Traveler type: do attitudes toward Rewinds and static photos differ depending on whether the participant identifies as a frequent traveler?
	\item GPS usage: do participants who are regular users of GPS navigation react differently to Rewinds and static photos compared to participants who don't use GPS navigation regularly?
\end{enumerate}

% TODO explain tables
\begin{table*}[t]
	\centering
	\caption{My caption}
	\small
	\label{within-subject-table}
	\begin{tabular}{c|c|c|c|c|c|c|c|c|}
		\cline{2-9}
		& \multicolumn{4}{c|}{Memorability} & \multicolumn{2}{c|}{Belivability} & \multicolumn{2}{c|}{Desirability} \\ \cline{2-9} 
		& \begin{tabular}[c]{@{}c@{}}video\\  location\end{tabular} & \multicolumn{2}{c|}{\begin{tabular}[c]{@{}c@{}}photo\\ location/trip\end{tabular}} & \begin{tabular}[c]{@{}c@{}}video\\ trip\end{tabular} & photo & video & photo & video \\ \hline
		\multicolumn{1}{|c|}{N} & 177 & \multicolumn{2}{c|}{179} & 177 & 86 & 128 & 179 & 177 \\ \hline
		\multicolumn{1}{|c|}{\%} & 96.61\% & \multicolumn{2}{c|}{49.72\%} & 53.67\% & 77.91\% & 79.69\% & 10.61\% & 14.69\% \\ \hline
		\multicolumn{1}{|c|}{p-val} & \multicolumn{2}{c|}{\begin{tabular}[c]{@{}c@{}}$\chi^2(1, 356) = 96.987$\\ p = 0.0001\end{tabular}} & \multicolumn{2}{c|}{\begin{tabular}[c]{@{}c@{}}$\chi^2(1, 356) = 0.41$\\ p = 0.2611\end{tabular}} & \multicolumn{2}{c|}{\begin{tabular}[c]{@{}c@{}}$\chi^2(1, 214) = 0.02$\\ p = 0.4434\end{tabular}} & \multicolumn{2}{c|}{\begin{tabular}[c]{@{}c@{}}$\chi^2(1, 356 ) = 0.995$\\ p = 0.1593\end{tabular}} \\ \hline
	\end{tabular}
\end{table*}

% 2nd table %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\section{Memorability}
Rewinds, and digitally-mediated memories in general, are meant to engage a person's memory. To find out what features make a digital memory memorable, we examined participant responses to whether or not they were able to remember the location shown in a photo (semantic memory) or recall the trip portrayed by a Rewind (episodic memory). Photo memorability is determined by whether or not the participant remembered the location depicted in a photo. Because a Rewind contains many locations, it would be unfair to consider the memorability of Rewind based on whether the participant remembers any locations in the Rewind. So, Rewind memorability is determined by whether or not the participant remembered and could distinguish the particular trip portrayed in a Rewind. In this section, we will look at differences in photo and Rewind memorability based on the five metrics discussed.

\subsubsection{Semantic Memory} 
Within-subject comparison in our study reveals that participants are significantly better at recalling semantic memory of a particular location with \textit{Rewind} videos than with static street-view images, $\chi^2(1, N=356) = 96.987$, $p <.01$. When being presented with a street-side image retrieved based on their location data, participants were able to recognize places visited 50\% of the time; yet recognition rate of locations reached 97\% when participants started playing accompanied Rewinds, with a significant increase of 47\%. It is perhaps not too surprising as cinematic media like Rewinds would provide users with more contextual meanings and retrival cues for semantic processing, as three participants (P4, P6 and P10) have once expressed their contentment after watching Rewinds by claiming ``I recognize the photo now!'' when their initial attempts with static images recognition were fruitless.

% shall we change photo/rewind video rememberance to location/trip rememberance? april - i define photo/rewind remembrance below. a photo and a rewind can have both location and trip remembrance. for instance from a photo you might remember the location but not the trip. 
\subsection{Memorability}
